\chapter{Serial Protocols and Safety State Machines}

\section{Why plain serial}

RUDRAv4 uses plain, line-oriented serial at the microcontroller boundary. This is simpler than embedding ROS inside microcontrollers and is easier to test with a serial monitor. Every protocol line is human-readable, newline-terminated, and short.

\section{Uno to NUC protocol}

The Uno emits:

\begin{lstlisting}[language={}]
J,select,ry,rx,ly,lx,guard
\end{lstlisting}

\begin{longtable}{p{0.16\linewidth}p{0.18\linewidth}p{0.56\linewidth}}
\toprule
Field & Type/range & Meaning\\
\midrule
\cmd{J} & literal & Packet type: joystick.\\
\cmd{select} & 0 or 1 & Latched manual-enable mode toggled by SELECT.\\
\cmd{ry} & 0--255-ish raw & Right stick Y raw value. Used only for optional guard chord logic if enabled.\\
\cmd{rx} & 0--255-ish raw & Right stick X raw value. Used for steering.\\
\cmd{ly} & 0--255-ish raw & Left stick Y raw value. Used for throttle.\\
\cmd{lx} & 0--255-ish raw & Left stick X raw value. Published but not presently used for driving.\\
\cmd{guard} & 0 or 1 & Latched obstacle guard enable toggled by START.\\
\bottomrule
\end{longtable}

The bridge accepts both the new seven-field packet and the older six-field form. If the guard field is absent, the bridge cannot update guard state from the controller latch.

\section{NUC to Teensy protocol}

The NUC emits:

\begin{lstlisting}[language={}]
D,left,right,enable
\end{lstlisting}

\begin{longtable}{p{0.16\linewidth}p{0.18\linewidth}p{0.56\linewidth}}
\toprule
Field & Type/range & Meaning\\
\midrule
\cmd{D} & literal & Packet type: drive command.\\
\cmd{left} & -127 to 127 & Left-side Sabertooth command after joystick mixing or \topic{/cmd_vel} conversion and obstacle filtering.\\
\cmd{right} & -127 to 127 & Right-side Sabertooth command.\\
\cmd{enable} & 0 or 1 & If 0, Teensy forces both sides to zero.\\
\bottomrule
\end{longtable}

The Teensy replies:

\begin{lstlisting}[language={}]
ACK,left,right,enable
ERR,bad_format
ERR,line_too_long
BOOT,RUDRA_TEENSY_SABERTOOTH_READY
BOOT,RUDRA_TEENSY_IMU_ODOM_READY
\end{lstlisting}

\section{Telemetry protocol}

The telemetry-capable Teensy firmware adds:

\begin{lstlisting}[language={}]
IMU,ax,ay,az,gx,gy,gz
ODOM,x,y,theta,linear_x,angular_z,left_vel,right_vel
\end{lstlisting}

The NUC parser treats these lines as telemetry and does not publish them as acknowledgements. All other Teensy lines go to \topic{/rudra/teensy_ack}. This allows the operator to see both control acknowledgement and richer telemetry without changing the base protocol.

\section{Timing and timeout}

The relevant timing constants are:

\begin{center}
\begin{tabular}{lcc}
\toprule
Loop & Period/timeout & Location\\
\midrule
Uno joystick print loop & 50 ms & Uno firmware\\
ROS bridge timer & 20 ms & Python node parameter\\
Teensy control loop & 20 ms & Teensy firmware\\
Teensy telemetry loop & 50 ms & telemetry firmware\\
Command timeout & 300 ms & ROS bridge and Teensy firmware\\
\bottomrule
\end{tabular}
\end{center}

The timeout design is layered. If the Uno disappears, the ROS bridge sends stop. If the NUC disappears, the Teensy times out and stops. If \topic{/cmd_vel} disappears, the \topic{/cmd_vel} bridge sends stop. No single software loop should be the only stop mechanism.

\begin{figure}[htbp]
\centering
\resizebox{0.98\linewidth}{!}{\begin{tikzpicture}[x=1cm,y=1cm]
\draw[->,thick] (0,0) -- (12,0) node[right]{time};
\foreach \x/\lab in {0/0,2/100ms,4/200ms,6/300ms,8/400ms,10/500ms} {
  \draw (\x,0.1) -- (\x,-0.1) node[below]{\lab};
}
\draw[rudraTeal,thick] (0,1.2) -- (5.8,1.2);
\draw[rudraAmber,thick] (5.8,1.2) -- (10.5,1.2);
\node[left] at (0,1.2) {commands};
\node[above] at (2.8,1.2) {valid \cmd{D,...}};
\node[above] at (8.2,1.2) {no update};
\draw[dashed] (6,0) -- (6,1.6);
\node[above] at (6,1.6) {timeout boundary};
\draw[rudraArrow] (6,0.65) -- (7.8,0.65) node[right]{Teensy forces target to zero};
\end{tikzpicture}}
\caption{Command timeout model. If fresh commands stop for more than about 300 ms, the Teensy disables drive.}
\end{figure}

\section{State machines}

The joystick bridge has a simple safety state machine:

\begin{enumerate}
  \item \textbf{No Uno / bad line}: send stop.
  \item \textbf{Valid packet but manual disabled}: send stop if configured.
  \item \textbf{Valid packet and enabled}: mix throttle/steering, apply obstacle guard, send drive.
  \item \textbf{Packet timeout}: send stop and publish zero \topic{/cmd_vel} preview.
\end{enumerate}

The Teensy state machine is even smaller:

\begin{enumerate}
  \item Parse latest \cmd{D,...} line.
  \item Clamp and store targets if enabled, otherwise zero targets.
  \item If command timeout expires, zero targets.
  \item Ramp actual commands toward targets.
  \item Write packet serial outputs to both Sabertooths.
\end{enumerate}

This split keeps high-level decisions on the NUC while ensuring the motor interface still has a local timeout if the NUC or USB link fails.
