\chapter{Launching, Operating, and Monitoring RUDRA}

\section{Pre-launch checklist}

Before launching robot nodes:

\begin{enumerate}
  \item Confirm motor power can be disconnected immediately.
  \item Confirm the wheels are lifted for the first test after any change.
  \item Confirm the Uno, Teensy, LiDAR, and DCDC-USB appear under \filep{/dev/serial/by-id} or expected USB listings.
  \item Close Arduino Serial Monitor or any terminal that owns a serial port.
  \item Confirm both Rudra and Aghora use \cmd{ROS_DOMAIN_ID=42} and \cmd{ROS_LOCALHOST_ONLY=0}.
  \item Source overlays in order: base ROS, external LiDAR workspace, then RUDRAv4.
\end{enumerate}

\section{Launch cheat sheet}

The normal operator entry point is the full launch file. It starts the robot-side graph: PS2 manual bridge, Teensy serial owner, LiDAR, obstacle guard, localization, voice, and optional helpers. Use this first unless a chapter explicitly asks for a lower-level commissioning launch.

\begin{lstlisting}[language=bash,caption={Default full robot launch.}]
cd /home/rudra/Projects/RUDRAv4
source /opt/ros/lyrical/setup.bash
source /home/rudra/ros2_ws/install/setup.bash
source install/setup.bash
ros2 launch rudra_base_bridge rudra_full.launch.py
\end{lstlisting}

\begin{longtable}{p{0.36\linewidth}p{0.54\linewidth}}
\toprule
Need & Launch option\\
\midrule
Normal full robot & none\\
Disable voice & \cmd{enable_voice:=false}\\
Show RViz on this machine & \cmd{enable_rviz:=true}\\
Enable NUC power diagnostics & \cmd{enable_dcdc_monitor:=true}\\
Start/check Ollama only & \cmd{enable_ollama:=true}\\
Let voice use Ollama & \cmd{enable_ollama:=true voice_use_llm_router:=true}\\
Use another LiDAR path & \cmd{serial_port:=/dev/serial/by-id/<actual-lidar-id>}\\
\bottomrule
\end{longtable}

\begin{lstlisting}[language=bash,caption={Common full-launch variants.}]
ros2 launch rudra_base_bridge rudra_full.launch.py enable_voice:=false
ros2 launch rudra_base_bridge rudra_full.launch.py enable_rviz:=true
ros2 launch rudra_base_bridge rudra_full.launch.py enable_dcdc_monitor:=true
ros2 launch rudra_base_bridge rudra_full.launch.py \
  enable_ollama:=true \
  voice_use_llm_router:=true
ros2 launch rudra_base_bridge rudra_full.launch.py \
  serial_port:=/dev/serial/by-id/<actual-lidar-id>
\end{lstlisting}

\begin{operatorbox}{One launch, optional toggles}
Do not run \filep{scripts/start_ollama.sh} before the full launch during normal operation. When \cmd{enable_ollama:=true}, \filep{rudra_full.launch.py} invokes that helper for you. The helper script remains useful for one-time diagnostics and model downloads.
\end{operatorbox}

\section{Mode selection}

\begin{longtable}{p{0.24\linewidth}p{0.34\linewidth}p{0.32\linewidth}}
\toprule
Mode & Use when & Command\\
\midrule
Full robot & Normal integrated bringup: PS2, Teensy, LiDAR, obstacle guard, localization, and voice & \cmd{ros2 launch rudra_base_bridge rudra_full.launch.py}\\
Full robot + Ollama & Full bringup with voice LLM routing enabled & \cmd{ros2 launch rudra_base_bridge rudra_full.launch.py voice_use_llm_router:=true enable_ollama:=true}\\
Full robot + DCDC & Normal NUC-mounted robot bringup with power diagnostics & \cmd{ros2 launch rudra_base_bridge rudra_full.launch.py enable_dcdc_monitor:=true}\\
PS2 only & Basic manual drive without LiDAR driver & \cmd{bash scripts/run_ps2_bridge.sh}\\
PS2 + LiDAR & Manual driving with \topic{/scan} and obstacle guard & \cmd{bash scripts/run_ps2_lidar_bridge.sh}\\
PS2 + LiDAR + localization & Full development bringup with IMU/wheel odom and EKF & \cmd{bash scripts/run_ps2_lidar_localization.sh}\\
\pkg{rudra_voice} & Voice command interface with P610, Vosk STT, optional Ollama, and guarded voice motion output & \cmd{ros2 launch rudra_voice rudra_voice.launch.py}\\
\topic{/cmd_vel} bridge & Keyboard teleop, Nav2, autonomy, or external command publisher & \cmd{bash scripts/run_cmd_vel_bridge.sh}\\
LiDAR only & Sensor and RViz debug without drive bridge & \cmd{bash scripts/run_lidar.sh}\\
DCDC monitor only & Power rail debug & \cmd{ros2 launch rudra_base_bridge dcdc_usb_monitor.launch.py}\\
\bottomrule
\end{longtable}

\begin{safetybox}{One Teensy owner only}
Do not run \pkg{ps2_uno_to_teensy} and \pkg{cmd_vel_to_teensy} at the same time. Both write to the same Teensy serial port. The full launch uses \pkg{ps2_uno_to_teensy} as the single Teensy owner. Voice motion enters that same bridge through \topic{/cmd_vel_safe} only when PS2/manual input is idle.
\end{safetybox}

\section{Full bringup sequence}

The clean operator command on \cmd{rudra} is:

\begin{lstlisting}[language=bash]
cd /home/rudra/Projects/RUDRAv4
source /opt/ros/lyrical/setup.bash
source /home/rudra/ros2_ws/install/setup.bash
source install/setup.bash
ros2 launch rudra_base_bridge rudra_full.launch.py
\end{lstlisting}

With the DCDC-USB powering the NUC, enable power diagnostics:

\begin{lstlisting}[language=bash]
ros2 launch rudra_base_bridge rudra_full.launch.py \
  enable_dcdc_monitor:=true
\end{lstlisting}

If Ollama is installed and the natural-language router should be checked before launch, add:

\begin{lstlisting}[language=bash]
ros2 launch rudra_base_bridge rudra_full.launch.py \
  enable_dcdc_monitor:=true \
  enable_ollama:=true \
  voice_use_llm_router:=true
\end{lstlisting}

The full launch starts these robot-side pieces:

\begin{itemize}
  \item \pkg{ps2_uno_to_teensy}: reads Uno PS2 packets, writes the only Teensy drive serial stream, publishes telemetry, and applies LiDAR obstacle avoidance.
  \item \pkg{ydlidar_ros2_driver_node}: publishes \topic{/scan} in frame \cmd{laser}.
  \item Static transforms for \cmd{base_link -> laser} and \cmd{base_link -> imu_link}.
  \item \pkg{robot_localization} EKF and \pkg{odom_tf_broadcaster}.
  \item \pkg{rudra_voice}: Vosk STT, deterministic intent parser, Piper/eSpeak speech output, and \pkg{command_guard_node}.
  \item Optional \pkg{dcdc_usb_monitor}, optional Ollama helper, and optional RViz.
\end{itemize}

Voice motion flow in the full launch is:

\begin{lstlisting}
voice_node -> /cmd_vel_voice_request -> command_guard_node -> /cmd_vel_safe -> ps2_uno_to_teensy -> Teensy
\end{lstlisting}

This means \topic{/cmd_vel_safe} is not merely a debug topic. In the full launch, it is the guarded voice input to the single Teensy bridge. PS2 input still wins: if the sticks or SELECT are active, voice motion is ignored until manual input is idle.

By default, the full launch keeps \pkg{use\_llm\_router} off. That makes the voice stack stick to deterministic commands like \cmd{status}, \cmd{stop}, and \cmd{move forward} even when Ollama is not running. \cmd{enable_ollama:=true} starts or checks the Ollama service. \cmd{voice_use_llm_router:=true} is the separate switch that lets voice send uncertain phrases to Ollama.

On Aghora:

\begin{lstlisting}[language=bash]
source /opt/ros/lyrical/setup.bash
export ROS_DOMAIN_ID=42
export ROS_LOCALHOST_ONLY=0
cd /home/rudra/Projects/RUDRAv4
source install/setup.bash
ros2 launch rudra_base_bridge localization_view.launch.py
\end{lstlisting}

\section{Monitoring commands}

Use a separate terminal for monitors:

\begin{lstlisting}[language=bash]
ros2 node list
ros2 topic list
ros2 topic echo /rudra/ps2_raw
ros2 topic echo /rudra/drive_cmd
ros2 topic echo /rudra/teensy_ack
ros2 topic echo /rudra/obstacle_guard
ros2 topic echo --once /scan
ros2 topic echo --once /rudra/imu/raw
ros2 topic echo --once /rudra/wheel_odom
ros2 topic echo --once /odometry/filtered
ros2 topic echo /tf
ros2 topic echo /rudra/nuc_battery
ros2 topic echo /diagnostics
ros2 topic echo /rudra_voice/transcript
ros2 topic echo /rudra_voice/intent
ros2 topic echo /rudra_voice/reply
ros2 topic echo /cmd_vel_voice_request
ros2 topic echo /cmd_vel_safe
\end{lstlisting}

\section{What good looks like}

\begin{longtable}{p{0.33\linewidth}p{0.57\linewidth}}
\toprule
Observation & Interpretation\\
\midrule
\topic{/rudra/ps2_raw} changes when sticks move & Uno and PS2 controller are reaching ROS.\\
\topic{/rudra/drive_cmd} shows bounded \cmd{D,...} lines & Bridge is computing final motor packets.\\
\topic{/rudra/teensy_ack} echoes \cmd{ACK,...} & Teensy is receiving drive commands.\\
\topic{/scan} publishes with frame \cmd{laser} & YDLIDAR driver is healthy enough for RViz and guard.\\
\topic{/rudra/obstacle_guard} changes when an obstacle is placed ahead & Scan reaches guard and forward scale logic is active.\\
\topic{/rudra/imu/raw} and \topic{/rudra/wheel_odom} publish & Telemetry firmware and parser are active.\\
\topic{/odometry/filtered} publishes & EKF has inputs and is alive.\\
\topic{/tf} includes \cmd{odom -> base_link} & RViz can place odom, scan, and robot body in the same tree.\\
\topic{/rudra/nuc_battery} publishes & DCDC-USB monitor sees the NUC power rail.\\
\topic{/diagnostics} contains \cmd{rudra_nuc_dcdc} & NUC input voltage, output range, and sag warnings are visible.\\
\topic{/rudra_voice/transcript} publishes after \cmd{Hey RUDRA} phrases & P610 audio, Vosk, and wake phrase processing are active.\\
\topic{/cmd_vel_safe} publishes a short pulse then zero & Voice guard accepted a bounded motion request and timed it out safely.\\
RViz displays scan and marker & DDS domain, TF, and display settings are correct.\\
\bottomrule
\end{longtable}

\section{RViz views}

The package includes two RViz views:

\begin{longtable}{p{0.28\linewidth}p{0.62\linewidth}}
\toprule
View & Purpose\\
\midrule
\cmd{lidar_view.launch.py} & Displays \topic{/scan} and the obstacle marker with fixed frame \cmd{base_link}. Use it for sensor and guard debugging.\\
\cmd{localization_view.launch.py} & Displays raw wheel odom, fused odom, scan, obstacle marker, and TF with fixed frame \cmd{odom}. Use it for localization and navigation readiness checks.\\
\bottomrule
\end{longtable}

RViz does not start the robot. It is a viewer. The robot-side launch must already be publishing the topics that the view consumes.

\section{Shutdown}

A conservative shutdown order is:

\begin{enumerate}
  \item Release sticks or stop the \topic{/cmd_vel} publisher.
  \item Stop the drive launch with Ctrl-C.
  \item Stop LiDAR if needed:
\begin{lstlisting}[language=bash]
bash scripts/stop_lidar.sh
\end{lstlisting}
  \item Remove Sabertooth motor power.
  \item Shut down the NUC cleanly:
\begin{lstlisting}[language=bash]
sudo shutdown now
\end{lstlisting}
\end{enumerate}

If the LiDAR still spins, it is still receiving USB power. Remove USB power physically.
